---
title: "Homework 2"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

**Name:** Aniketh Kalagara

**UT EID:** ak54432

**Github Repo:** https://github.com/anikethk1/homework2_sds315

```{r}
library(ggplot2)
library(tidyverse)
profs <- read.csv("profs.csv")
bike_share <- read.csv("bikeshare.csv")
```

# **Question 1:**
```{r}
ggplot(profs) + 
  geom_histogram(aes(x = eval)) +
  labs(title = "Histogram of Course Evaluations", x = "2-5 Rating Scale", y = "Frequency")
```

The mean evaluation rating is about 4.00 and the standard deviation is 0.55. The plot also has a negative skew. All of this information shows that most courses at UT Austin have a strong rating at about 4/5 and that the courses offered are of good use for the students. There is variation in course evaluations as with a variety of courses, some are bound to have a low rating, still with such a high average, it shows the strong performance of the university.

```{r}
ggplot(profs) +
  geom_boxplot(aes(x = eval, fill = native)) + 
  labs(title = "Boxplots of Course Evaluations", x = "2-5 Rating Scale", fill = "Native English Speaker") +
  theme(axis.text.y = element_blank()) +
  facet_wrap(~native, nrow = 2)
```

The mean rating of professors who are native english speakers is 4.0 compared to non-native english speakers being 3.69. Both plots have an IQR of 0.7 The reason for this variation could possibly be due to ease in understanding and relating to the professor. This allows for the material to be presented better by a native english speaker as most if not all students at UT Austin speak english. With an easier to understand course, the ratings are bound to be more. The same IQRs shows how even the middle 50% of values are around their mean value and that there is a clear and stark difference between native and non-native professors and their course performance, making it an important factor for course evaluations.

```{r}
ggplot(profs) + 
  geom_histogram(aes(x = eval, fill = gender)) + 
  labs(title = "Histogram of Course Evaluations", x = "2-5 Rating Scale", y = 'Frequency', fill = "Professor Gender") +
  facet_wrap(~gender, nrow=2)
```

Male professors have a slightly higher mean rating at 4.07 compared to female professors at 3.90. Both graphs have a greater negative skew and the standard deviations of both genders are very similar, 0.539 and 0.557. A factor that could affect the female professor ratings are the fact that there are significantly less female professors, 195 female to 268 male to be exact. If there were more female professors, there could be a possibility of females having a higher or lower rating. Otherwise, there are similarities across summary statistics, simply with males having a slightly higher average rating. This does not create the most conclusive evidence of males being better professors, but according to the given data at UT Austin on average, students do prefer male professors. Although, this factor isn’t decisive in a course’s evaluation due to the imbalance in number of professors per gender and the extreme similarities in all statistics and skew.

```{r}
ggplot(profs) +
  geom_point(aes(x = beauty,y = eval)) +
  labs(title = "Scatter Plot of Professor Attractiveness vs Course Evaluation",  x= "Professor Attractiveness", y = "Course  Rating")
```

As seen by the scatter plot and by the correlation of 0.19, it can be concluded that there is a very weak positive correlation between professor attractiveness and course performance. There might be a very slight factor at play that students might be more lenient or like a professor who looks better, but due to the weakness of correlation, it is not a defining factor towards course performance.

# **Question 2**

```{r}
bikeshare_tib <- read_csv("bikeshare.csv")

bikeshare_tib <- mutate(bikeshare_tib, workingday = replace(workingday, workingday == 1, "Working Day"))
bikeshare_tib <- mutate(bikeshare_tib, workingday = replace(workingday, workingday == 0, "Non-working Day"))

bikeshare_byhours <- group_by(bikeshare_tib, hr)
bikeshare_byhour_average <- summarize(bikeshare_byhours, average_rentals = mean(total))
bikeshare_byhour_average <- as.data.frame(bikeshare_byhour_average)

ggplot(bikeshare_byhour_average) + 
  geom_line(aes(x = hr, y = average_rentals)) +
  labs(title = "Line Graph of Hourly Bike Rentals across Hours of the Day",  x = "Hour of the Day (hr)", y = "Average Number of Bike Rentals")
```

This line graph shows the changes in the number of average bike rentals per hour across two years in Washington DC. The peaks can be seen to occur around 8 and 17, 18. This indicates 7 AM and 4-5 PM as peak times for bike sharing to be used. This can be directly correlated to people going to and from work or school. The reason there are drops is those are the peak traffic times during the course of the day, and most individuals would avoid traffic through these bike riding features while also being environmentally friendly in the process. To take home, most riders use bikes during peak traffic hours to enable ease of transportation to and from work or school. 

```{r}
bikeshare_byhour_workingday <- group_by(bikeshare_tib, hr, workingday)
bikeshare_byhour_workingday_average <- summarize(bikeshare_byhour_workingday, average_rentals = mean(total))
bikeshare_byhour_workingday_average <- as.data.frame(bikeshare_byhour_workingday_average)

ggplot(bikeshare_byhour_workingday_average) +
  geom_line(aes(x = hr, y = average_rentals, color = workingday)) +
  labs(title = "Line Graph of Hourly Bike Rentals across Hours of the Day",  x = "Hour of the Day (hr)", y = "Average Number of Bike Rentals", color = "Working Day") +
  facet_wrap(~workingday)
```

This line graph is faceted by whether the day is a working day or not. As mentioned before, working hours are probably what effects the peaks in the graph which is seen by the peaks at the same times of 8, 17, and 18 hours. Meanwhile, on the non-working days, the graph is a lot more balanced in the sense of there being most peaks during hours which individuals are most likely to be awake and actually able enough to use a bike, rather than in the early morning or night when they are tired, from 12 to 16. To take home, these graphs show the importance of a day being a working day and not a working day accordingly affecting the hours that riders use bikes, making it a defining factor.

```{r}
bikeshare_9am <- filter(bikeshare_tib, hr == 8)
bikeshare_9am <- group_by(bikeshare_9am, weathersit, workingday)
bikeshare_9am <- summarize(bikeshare_9am, average_rentals = mean(total))
bikeshare_9am <- as.data.frame(bikeshare_9am)

bikeshare_9am <- mutate(bikeshare_9am, workingday = replace(workingday, workingday == "Yes", "Working Day"))
bikeshare_9am <- mutate(bikeshare_9am, workingday = replace(workingday, workingday == "No", "Non-working Day"))

ggplot(bikeshare_9am) +
  geom_bar(aes(x = weathersit, y = average_rentals, fill = workingday), stat="identity") +
  labs(title = "Bar Plot of Average Bike Ridership at 9AM by Weather Situation", x = "Weather Situation", y = "Average Number of Bike Rentals", fill = "Working Day") +
  facet_wrap(~workingday)
```

These bar plots show the number of bike rentals according to weather condition. The trend is pretty similar across working and non-working days, but specific to the weather conditions, that consistency is important to identify. The average number of bike rentals is very similar for clear or partly cloudy and misty weather. This is due to the ease of vision and safety while travelling in such conditions. Snowy or rainy conditions cause the number of bikes used to drop considerably due to obvious safety issues with travelling in such weather. To take home, the weather is extremely important in measuring riders usage of bikes as safety is of utmost importance when using said mode of transport. 

# **Question 3**

```{r}
capmetro <- read.csv("capmetro_UT.csv")
capmetro = mutate(capmetro,
 day_of_week = factor(day_of_week,
 levels=c("Mon", "Tue", "Wed","Thu", "Fri", "Sat", "Sun")),
 month = factor(month,
 levels=c("Sep", "Oct","Nov")))

capmetro_tib <- as_tibble(capmetro)
capmetro_tib <- group_by(capmetro_tib, day_of_week, hour_of_day, month)
capmetro_tib <- summarize(capmetro_tib, average_boardings = mean(boarding))
capmetro_df <- as.data.frame(capmetro_tib)

ggplot(capmetro_df) +
  geom_line(aes(x = hour_of_day, y = average_boardings, color = month)) +
  labs(title = "Line Graph of Average Capital Metro Boardings per Hour by Day of Week", x = "Hour of Day", y = "Average Boardings", color = "Month") +
  facet_wrap(~day_of_week)
```

The figure shows line graphs of the average number of Capital Metro bus boardings across the seven days of the week based on the hour of the day. Each graph has three lines representing the trend across the months of September, October and November. The hour of peak boarding across weekdays is fairly similar, with it being around the span of 15-17 or 3-5PM, this remained mostly the same across all months. During weekends it is a lot harder to tell where the peaks are due to a much smaller amount of bus riders in the first place, but there is a slight increase in bus travelling on Sundays towards the later half while on Saturday it mostly remains consistent from 10-19 or 10AM-7PM. Mondays in September could have probably had a drop due to the Labor Day holiday which would have probably caused the usage of the campus bus to be extremely small, bringing the average down greatly. Wednesday, Thursday, and Friday could have possibly dropped in November due to the Thanksgiving weekend where a lot of students head back home. 

```{r}
ggplot(capmetro) +
  geom_point(aes(x = temperature, y = boarding, color = weekend)) +
  labs(title = "Scatter Plot of Temperature vs Capital Metro Boardings by Hour of Day", x = "Temperature (F)", y = "Boardings", color = "Weekday or Weekend") +
  facet_wrap(~hour_of_day)
```

The figure shows various faceted scatter plots according to the hour of day in which each plot measures the number of boardings based on the temperature on that day. Each scatter plot has it's points divided by weather it is a weekday or weekend. If hour of day and weekend status is kept constant, it can be inferred that temperature isn't a big factor in UT students entering the bus. There are more points towards the right of each of the figures and that is because Austin as an area is quite warm and higher temperatures are more common. There is no real correlation between temperature and boarding, justified by the weak correlation of 0.19, there might be a slight role to play, but due to the smaller amount of values at colder temperatures, the correlation is not high enough to judge that temperature plays a role in UT students using the bus. 

\newpage

# **Question 4**

```{r}
billboard <- read_csv("billboard.csv")

top10 <- group_by(billboard, performer, song)
top10 <- summarize(top10, count = n())
top10 <- arrange(top10, desc(count))
print(top10)
```

This table shows the 10 most popular songs based on the number of weeks they were on the Billboard Top 100, the most being Imagine Dragon's Radioactive at 87 weeks on the billboard. 

```{r}
unique_songs <- filter(billboard, year > 1958 & year < 2021)
unique_songs <- group_by(unique_songs, year)
unique_songs <- summarize(unique_songs, num_songs = length(unique(song)))

unique_songs_df <- as.data.frame(unique_songs)

ggplot(unique_songs_df) +
  geom_line(aes(x = year, y = num_songs)) +
  labs(title = "Line Graph of Year vs Number of Unique Songs on Billboard Top 100", x = "Year", y = "Number of Songs")
```

The line graph showcases the number of unique songs on the Billboard Top 100 in an attempt to measure musical diversity over the years. The graph has a fluctuating nature but there is an overall clear trend seen. There was a peak in unique songs on the billboard around 1965 after going on a steady decrease to the early 2000s before rapidly moving back up as it keeps increasing until the year 2020. This could possibly represent a transition between a generation of music. During the dip, old artists were retiring and new artists were up and coming, so mainstream music would have fallen in creating new music trends and as new artists started to gain popularity, so did the uniqueness on the Billboard Top 100. 

```{r}
ten_week_hits <- filter(top10, count >= 10)
ten_week_hits <- group_by(ten_week_hits, performer)
ten_week_hits <- summarize(ten_week_hits, num_hits = n())
ten_week_hits <- filter(ten_week_hits, num_hits >= 30)

ten_week_hits_df <- as.data.frame(ten_week_hits)

ggplot(ten_week_hits_df) +
  geom_bar(aes(x = performer, y = num_hits), stat="identity") +
  coord_flip() +
  labs(title = "Bar Graph of Artists with 30+ Ten-Week Hits", x = "Artist Name", y = "Number of Ten-Week Hits")
```

The graph shows the 19 artists who have 30 or more ten-week hits, songs which have been on the Billboard Top 100 for 10 or more weeks. As seen, Elton John has by far the largest amount of songs at 50+ songs which have been ten-week hits. 









